---
title: "first rasch model"
author: "Iris Zhong"
date: "1/25/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Getting started ---------------------------------------------------------
## this will be our maiden voyage estimating IRT models
## the main goal is to get oriented with the key output 
## that we get from applications of these models to item response data


```{r}
resp1 <- read.table("https://github.com/ben-domingue/252L/raw/master/data/emp-rasch.txt", header=FALSE)
```


## fit the rasch model


```{r}
library(mirt) # might have to run install.packages("mirt")
m1 <- mirt(resp1, 1, itemtype = "Rasch")
m1 #here is the object containing the estimated rasch model. it contains lots of stuff, we're just seeing a quick summary here
```


##it has plot methods attached that will generate item response functions (or trace lines, as they are called here)

```{r}
plot(m1, type = "trace") ## which is the easiest item? the most difficult item?
```

plot from -6 to 6


## we can use the below to get item parameters

```{r}
coef(m1)
```



##that was unwieldy, here is a better way of getting that output


```{r}
get_coef <- function(mod) {
  co <- coef(mod)
  co <- co[-length(co)]#why do i get rid of this last bit?
  do.call("rbind", co)
}
coef<-get_coef(m1) #first column is alpha (note that they don't vary), second column is 'difficulty', ignore third and fourth
coef
```


## in particular, i would look over this closely vis-a-vis the relevant part of the mirt manual:
## Rasch Only one intercept estimated, and the latent variance of
##      theta is freely estimated. If the data have more than two
##      categories then a partial credit model is used instead (see
##      'gpcm' below).
##           P(x = 1|theta, d) = \frac{1}{1 + exp(-(theta + d))}      

##note that there is something different when we compare "our" version of the Rasch model to the mirt version.
##It's very important that you note this difference!
##so, be able to make sure you can explain this!



```{r}
plot(colMeans(resp1,na.rm=TRUE),coef[,2],xlab="p-values",ylab="raw mirt estimates")
```

higher p-value ~ rasch item easiness (not difficulty!)
d = -b (easiness = -difficulty)

## here is a fun way of looking at comparing the estimated icc to empirical data


```{r}
itemfit(m1, empirical.plot = 48)
```

dot: actual observation with that theta
assess the accuracy of the rasch model

